Semi-parametric generalized linear model for correlated binary data

### 2021-03-07 nuweb/SPGLM.tex
Monitor log-likelihood during convergence

\begin{flushleft} \small\label{scrap1}\raggedright\small
\NWtarget{nuweb1}{} \verb@"../R/SPGLM.R"@\nobreak\ {\footnotesize {1}}$\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@#' Fit semi-parametric GLM@\\
\mbox{}\verb@#'@\\
\mbox{}\verb@#'@{\tt @}\verb@rdname spglm@\\
\mbox{}\verb@#'@{\tt @}\verb@param formula a one-sided formula of the form \code{cbind(r, s) ~ predictors} where \code{r} and \code{s} give the number of responses and non-responses within each cluster, respectively (so the cluster size is \code{r+s}), and \code{predictors} describes the covariates.@\\
\mbox{}\verb@#'@{\tt @}\verb@param data  an optional matrix or data frame containing the variables in the formula \code{formula}. By default the variables are taken from \code{environment(formula).}@\\
\mbox{}\verb@#'@{\tt @}\verb@param subset  an optional vector specifying a subset of observations to be used.@\\
\mbox{}\verb@#'@{\tt @}\verb@param weight  an optional vector specifying observation weights.@\\
\mbox{}\verb@#'@{\tt @}\verb@param link      a link function for the mean.@\\
\mbox{}\verb@#'@{\tt @}\verb@param mu0       an optional numeric value constraining the mean of the baseline distribution@\\
\mbox{}\verb@#'@{\tt @}\verb@param control a list with parameters controlling the algorithm.@\\
\mbox{}\verb@#'@{\tt @}\verb@return an object of class \code{spglm} with the fitted model.@\\
\mbox{}\verb@#'@{\tt @}\verb@export@\\
\mbox{}\verb@#' @{\tt @}\verb@importFrom stats terms model.matrix@\\
\mbox{}\verb@@\\
\mbox{}\verb@spglm <- function(formula, data, subset, weights, offset, link="logit", mu0=NULL, @\\
\mbox{}\verb@                  control=list(eps=0.001, maxit=100), ...){@\\
\mbox{}\verb@@\\
\mbox{}\verb@    @\hbox{$\langle\,${\itshape Create model matrix from formula and data}\nobreak\ {\footnotesize \NWlink{nuweb2}{2}}$\,\rangle$}\verb@@\\
\mbox{}\verb@    @\hbox{$\langle\,${\itshape Fit model}\nobreak\ {\footnotesize \NWlink{nuweb3b}{3b}}$\,\rangle$}\verb@@\\
\mbox{}\verb@@\\
\mbox{}\verb@    mt <- attr(mf, "terms")@\\
\mbox{}\verb@    res <- list(coefficients = betas, f0=referencef0, mu0=mu0, niter = iter, loglik=mod$llik,@\\
\mbox{}\verb@                link = link, call = mc, terms = mt,@\\
\mbox{}\verb@                xlevels = .getXlevels(mt, mf),@\\
\mbox{}\verb@                data_object=list(model_matrix=mm, resp=Y, n=rowSums(Y), weights=weights, offset=offset))@\\
\mbox{}\verb@    class(res) <- "spglm"@\\
\mbox{}\verb@    res@\\
\mbox{}\verb@@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtFileDefBy\ \NWlink{nuweb1}{1}\NWlink{nuweb3a}{, 3a}.

@@

\begin{flushleft} \small\label{scrap4}\raggedright\small
\NWtarget{nuweb3b}{} $\langle\,${\itshape Fit model}\nobreak\ {\footnotesize {3b}}$\,\rangle\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@ @\hbox{$\langle\,${\itshape Set starting values}\nobreak\ {\footnotesize \NWlink{nuweb6}{6}}$\,\rangle$}\verb@@\\
\mbox{}\verb@ @\hbox{$\langle\,${\itshape Set up for E-step}\nobreak\ {\footnotesize \NWlink{nuweb4}{4}}$\,\rangle$}\verb@@\\
\mbox{}\verb@ @\\
\mbox{}\verb@ iter <- 0@\\
\mbox{}\verb@ difference <- 100@\\
\mbox{}\verb@ while (iter < control$maxit & difference > control$eps) {@\\
\mbox{}\verb@    iter <- iter + 1@\\
\mbox{}\verb@    referencef0Pre <- referencef0@\\
\mbox{}\verb@    betasPre <- betas@\\
\mbox{}\verb@    @\\
\mbox{}\verb@    @\hbox{$\langle\,${\itshape E-step}\nobreak\ {\footnotesize \NWlink{nuweb5a}{5a}}$\,\rangle$}\verb@@\\
\mbox{}\verb@    @\hbox{$\langle\,${\itshape M-step}\nobreak\ {\footnotesize \NWlink{nuweb5b}{5b}}$\,\rangle$}\verb@@\\
\mbox{}\verb@    @\\
\mbox{}\verb@    difference <- sum(abs(referencef0Pre - referencef0)) + sum(abs(betasPre - betas))@\\
\mbox{}\verb@  }@\\
\mbox{}\verb@  @\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtMacroRefIn\ \NWlink{nuweb1}{1}.

@@

\begin{flushleft} \small\label{scrap5}\raggedright\small
\NWtarget{nuweb4}{} $\langle\,${\itshape Set up for E-step}\nobreak\ {\footnotesize {4}}$\,\rangle\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@  # replace each cluster with N+1 clusters of size N@\\
\mbox{}\verb@  new <- cbind(y=0:N)@\\
\mbox{}\verb@  rep_idx <- rep(1:nrow(Y), each=nrow(new))@\\
\mbox{}\verb@  Y2 <- cbind(1:nrow(Y), Y)[rep_idx,]@\\
\mbox{}\verb@  colnames(Y2) <- c("i", "resp","nonresp")@\\
\mbox{}\verb@@\\
\mbox{}\verb@  rep_idx2 <- rep(1:nrow(new), times=nrow(Y))@\\
\mbox{}\verb@  Ycomb <- cbind(Y2, new[rep_idx2, ,drop=FALSE])@\\
\mbox{}\verb@  # select possible combinations@\\
\mbox{}\verb@  possible <- (Ycomb[,"y"] >= Ycomb[,"resp"]) & (N-Ycomb[,"y"] >= Ycomb[,"nonresp"])@\\
\mbox{}\verb@  Ycomb <- Ycomb[possible,]@\\
\mbox{}\verb@ @\\
\mbox{}\verb@  mm2 <- mm[Ycomb[,"i"], ,drop=FALSE]@\\
\mbox{}\verb@  weights2 <- weights[Ycomb[,"i"]]@\\
\mbox{}\verb@  offset2 <- offset[Ycomb[,"i"]]@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtMacroRefIn\ \NWlink{nuweb3b}{3b}.

@@

\begin{flushleft} \small\label{scrap6}\raggedright\small
\NWtarget{nuweb5a}{} $\langle\,${\itshape E-step}\nobreak\ {\footnotesize {5a}}$\,\rangle\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@  # convert fTiltMatrix to long vector, watching out for potentially different support@\\
\mbox{}\verb@  y_idx <- match(Ycomb[,"y"], spt)@\\
\mbox{}\verb@  fTilt2 <- ifelse(!is.na(y_idx), fTiltMatrix[cbind(Ycomb[,"i"], y_idx)], 0)@\\
\mbox{}\verb@  @\\
\mbox{}\verb@  # numerator of weights@\\
\mbox{}\verb@  pp_num <- choose(n=Ycomb[,"y"], k=Ycomb[,"resp"]) * @\\
\mbox{}\verb@           choose(n=N-Ycomb[,"y"], k=Ycomb[,"nonresp"]) * @\\
\mbox{}\verb@           fTilt2@\\
\mbox{}\verb@  # denominator @\\
\mbox{}\verb@  pp_denom <- tapply(pp_num, list(i=Ycomb[,"i"]), sum, simplify=TRUE)@\\
\mbox{}\verb@  pp <- c(pp_num / pp_denom[Ycomb[,"i"]])@\\
\mbox{}\verb@  if (!is.null(weights)) pp <- weights2 * pp@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtMacroRefIn\ \NWlink{nuweb3b}{3b}.

@@

\begin{flushleft} \small\label{scrap7}\raggedright\small
\NWtarget{nuweb5b}{} $\langle\,${\itshape M-step}\nobreak\ {\footnotesize {5b}}$\,\rangle\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@  gldrmControl0 <- gldrm.control(returnfTiltMatrix = TRUE, returnf0ScoreInfo = FALSE, print=FALSE,@\\
\mbox{}\verb@                                betaStart = betasPre, f0Start = referencef0Pre)@\\
\mbox{}\verb@@\\
\mbox{}\verb@  mod <- gldrmFit(x = mm2, y=Ycomb[,"y"]/N, linkfun=link$linkfun, linkinv = link$linkinv,@\\
\mbox{}\verb@                  mu.eta = link$mu.eta, mu0 = mu0, offset = offset2, weights = pp, @\\
\mbox{}\verb@                  gldrmControl = gldrmControl0,  thetaControl=theta.control(),@\\
\mbox{}\verb@                  betaControl=beta.control(), f0Control=f0.control())@\\
\mbox{}\verb@                  @\\
\mbox{}\verb@  fTiltMatrix <- mod$fTiltMatrix@\\
\mbox{}\verb@  betas <- mod$beta@\\
\mbox{}\verb@  referencef0 <- mod$f0@\\
\mbox{}\verb@  spt <- round(mod$spt * N)@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtMacroRefIn\ \NWlink{nuweb3b}{3b}.

@@

\begin{flushleft} \small\label{scrap8}\raggedright\small
\NWtarget{nuweb6}{} $\langle\,${\itshape Set starting values}\nobreak\ {\footnotesize {6}}$\,\rangle\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@  N <- max(rowSums(Y))@\\
\mbox{}\verb@  nobs <- nrow(mm)@\\
\mbox{}\verb@  p <- ncol(mm)@\\
\mbox{}\verb@  @\\
\mbox{}\verb@  betas <- rep(0, times= p)@\\
\mbox{}\verb@@\\
\mbox{}\verb@  pooled <- CBData(data.frame(Trt = "All", NResp = Y[,1], ClusterSize = rowSums(Y), Freq=ceiling(weights)), @\\
\mbox{}\verb@                    trt="Trt", clustersize="ClusterSize", nresp="NResp")@\\
\mbox{}\verb@  est <- mc.est(pooled)@\\
\mbox{}\verb@  referencef0 <- est$Prob[est$ClusterSize == N]@\\
\mbox{}\verb@  # ensure all positive values@\\
\mbox{}\verb@  referencef0 <- (referencef0 + 1e-6)/(1+(N+1)*1e-6)@\\
\mbox{}\verb@  @\\
\mbox{}\verb@  fTiltMatrix <- matrix(rep(referencef0, times=nobs), byrow=TRUE, nrow=nobs, ncol=N+1)@\\
\mbox{}\verb@  spt <- 0:N@\\
\mbox{}\verb@@\\
\mbox{}\verb@  if (is.null(mu0))@\\
\mbox{}\verb@    mu0 <- weighted.mean(Y[,1]/rowSums(Y), weights)@\\
\mbox{}\verb@  @\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtMacroRefIn\ \NWlink{nuweb3b}{3b}.

>>>>

\begin{flushleft} \small\label{scrap1}\raggedright\small
\NWtarget{nuweb1}{} \verb@"../R/SPGLM.R"@\nobreak\ {\footnotesize {1}}$\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@#' Fit semi-parametric GLM@\\
\mbox{}\verb@#'@\\
\mbox{}\verb@#'@{\tt @}\verb@rdname spglm@\\
\mbox{}\verb@#'@{\tt @}\verb@param formula a one-sided formula of the form \code{cbind(r, s) ~ predictors} where \code{r} and \code{s} give the number of responses and non-responses within each cluster, respectively (so the cluster size is \code{r+s}), and \code{predictors} describes the covariates.@\\
\mbox{}\verb@#'@{\tt @}\verb@param data  an optional matrix or data frame containing the variables in the formula \code{formula}. By default the variables are taken from \code{environment(formula).}@\\
\mbox{}\verb@#'@{\tt @}\verb@param subset  an optional vector specifying a subset of observations to be used.@\\
\mbox{}\verb@#'@{\tt @}\verb@param weight  an optional vector specifying observation weights.@\\
\mbox{}\verb@#'@{\tt @}\verb@param link      a link function for the mean.@\\
\mbox{}\verb@#'@{\tt @}\verb@param mu0       an optional numeric value constraining the mean of the baseline distribution@\\
\mbox{}\verb@#'@{\tt @}\verb@param control a list with parameters controlling the algorithm.@\\
\mbox{}\verb@#'@{\tt @}\verb@return an object of class \code{spglm} with the fitted model.@\\
\mbox{}\verb@#'@{\tt @}\verb@export@\\
\mbox{}\verb@#' @{\tt @}\verb@importFrom stats terms model.matrix@\\
\mbox{}\verb@@\\
\mbox{}\verb@spglm <- function(formula, data, subset, weights, offset, link="logit", mu0=NULL, @\\
\mbox{}\verb@                  control=list(eps=0.001, maxit=100), ...){@\\
\mbox{}\verb@@\\
\mbox{}\verb@    @\hbox{$\langle\,${\itshape Create model matrix from formula and data}\nobreak\ {\footnotesize \NWlink{nuweb2}{2}}$\,\rangle$}\verb@@\\
\mbox{}\verb@    @\hbox{$\langle\,${\itshape Fit model}\nobreak\ {\footnotesize \NWlink{nuweb3b}{3b}}$\,\rangle$}\verb@@\\
\mbox{}\verb@@\\
\mbox{}\verb@    mt <- attr(mf, "terms")@\\
\mbox{}\verb@    res <- list(coefficients = betas, f0=referencef0, mu0=mu0, niter = iter, loglik=llik,@\\
\mbox{}\verb@                link = link, call = mc, terms = mt,@\\
\mbox{}\verb@                xlevels = .getXlevels(mt, mf),@\\
\mbox{}\verb@                data_object=list(model_matrix=mm, resp=Y, n=rowSums(Y), weights=weights, offset=offset))@\\
\mbox{}\verb@    class(res) <- "spglm"@\\
\mbox{}\verb@    res@\\
\mbox{}\verb@@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtFileDefBy\ \NWlink{nuweb1}{1}\NWlink{nuweb3a}{, 3a}.

@@

\begin{flushleft} \small\label{scrap4}\raggedright\small
\NWtarget{nuweb3b}{} $\langle\,${\itshape Fit model}\nobreak\ {\footnotesize {3b}}$\,\rangle\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@ @\hbox{$\langle\,${\itshape Set starting values}\nobreak\ {\footnotesize \NWlink{nuweb6b}{6b}}$\,\rangle$}\verb@@\\
\mbox{}\verb@ @\hbox{$\langle\,${\itshape Set up for E-step}\nobreak\ {\footnotesize \NWlink{nuweb4}{4}}$\,\rangle$}\verb@@\\
\mbox{}\verb@ @\\
\mbox{}\verb@ iter <- 0@\\
\mbox{}\verb@ difference <- 100@\\
\mbox{}\verb@ while (iter < control$maxit & difference > control$eps) {@\\
\mbox{}\verb@    iter <- iter + 1@\\
\mbox{}\verb@    referencef0Pre <- referencef0@\\
\mbox{}\verb@    betasPre <- betas@\\
\mbox{}\verb@    llikPre <- llik@\\
\mbox{}\verb@    @\\
\mbox{}\verb@    @\hbox{$\langle\,${\itshape E-step}\nobreak\ {\footnotesize \NWlink{nuweb5}{5}}$\,\rangle$}\verb@@\\
\mbox{}\verb@    @\hbox{$\langle\,${\itshape M-step}\nobreak\ {\footnotesize \NWlink{nuweb6a}{6a}}$\,\rangle$}\verb@@\\
\mbox{}\verb@    @\\
\mbox{}\verb@    difference <- abs(llik - llikPre)@\\
\mbox{}\verb@  }@\\
\mbox{}\verb@  @\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtMacroRefIn\ \NWlink{nuweb1}{1}.

@@

\begin{flushleft} \small\label{scrap5}\raggedright\small
\NWtarget{nuweb4}{} $\langle\,${\itshape Set up for E-step}\nobreak\ {\footnotesize {4}}$\,\rangle\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@  # replace each cluster with N+1 clusters of size N@\\
\mbox{}\verb@  new <- cbind(y=0:N)@\\
\mbox{}\verb@  rep_idx <- rep(1:nrow(Y), each=nrow(new))@\\
\mbox{}\verb@  Y2 <- cbind(1:nrow(Y), Y)[rep_idx,]@\\
\mbox{}\verb@  colnames(Y2) <- c("i", "resp","nonresp")@\\
\mbox{}\verb@@\\
\mbox{}\verb@  rep_idx2 <- rep(1:nrow(new), times=nrow(Y))@\\
\mbox{}\verb@  Ycomb <- cbind(Y2, new[rep_idx2, ,drop=FALSE])@\\
\mbox{}\verb@  # select possible combinations@\\
\mbox{}\verb@  possible <- (Ycomb[,"y"] >= Ycomb[,"resp"]) & (N-Ycomb[,"y"] >= Ycomb[,"nonresp"])@\\
\mbox{}\verb@  Ycomb <- Ycomb[possible,]@\\
\mbox{}\verb@  @\\
\mbox{}\verb@  obs_start <- match(1:nobs, Ycomb[,"i"])@\\
\mbox{}\verb@ @\\
\mbox{}\verb@  mm2 <- mm[Ycomb[,"i"], ,drop=FALSE]@\\
\mbox{}\verb@  weights2 <- weights[Ycomb[,"i"]]@\\
\mbox{}\verb@  offset2 <- offset[Ycomb[,"i"]]@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtMacroRefIn\ \NWlink{nuweb3b}{3b}.

@@

\begin{flushleft} \small\label{scrap6}\raggedright\small
\NWtarget{nuweb5}{} $\langle\,${\itshape E-step}\nobreak\ {\footnotesize {5}}$\,\rangle\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@  # convert fTiltMatrix to long vector, watching out for potentially different support@\\
\mbox{}\verb@  y_idx <- match(Ycomb[,"y"], spt)@\\
\mbox{}\verb@  fTilt2 <- ifelse(!is.na(y_idx), fTiltMatrix[cbind(Ycomb[,"i"], y_idx)], 0)@\\
\mbox{}\verb@  @\\
\mbox{}\verb@  # numerator of weights@\\
\mbox{}\verb@  pp_num <- choose(n=Ycomb[,"y"], k=Ycomb[,"resp"]) * @\\
\mbox{}\verb@           choose(n=N-Ycomb[,"y"], k=Ycomb[,"nonresp"]) * @\\
\mbox{}\verb@           fTilt2@\\
\mbox{}\verb@  # denominator @\\
\mbox{}\verb@  pp_denom <- tapply(pp_num, list(i=Ycomb[,"i"]), sum, simplify=TRUE)@\\
\mbox{}\verb@  pp <- c(pp_num / pp_denom[Ycomb[,"i"]])@\\
\mbox{}\verb@  if (!is.null(weights)) pp <- weights2 * pp@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtMacroRefIn\ \NWlink{nuweb3b}{3b}.

@@

\begin{flushleft} \small\label{scrap7}\raggedright\small
\NWtarget{nuweb6a}{} $\langle\,${\itshape M-step}\nobreak\ {\footnotesize {6a}}$\,\rangle\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@  gldrmControl0 <- gldrm.control(returnfTiltMatrix = TRUE, returnf0ScoreInfo = FALSE, print=FALSE,@\\
\mbox{}\verb@                                betaStart = betasPre, f0Start = referencef0Pre)@\\
\mbox{}\verb@@\\
\mbox{}\verb@  mod <- gldrmFit(x = mm2, y=Ycomb[,"y"]/N, linkfun=link$linkfun, linkinv = link$linkinv,@\\
\mbox{}\verb@                  mu.eta = link$mu.eta, mu0 = mu0, offset = offset2, weights = pp, @\\
\mbox{}\verb@                  gldrmControl = gldrmControl0,  thetaControl=theta.control(),@\\
\mbox{}\verb@                  betaControl=beta.control(), f0Control=f0.control())@\\
\mbox{}\verb@                  @\\
\mbox{}\verb@  fTiltMatrix <- mod$fTiltMatrix[obs_start,]@\\
\mbox{}\verb@  betas <- mod$beta@\\
\mbox{}\verb@  referencef0 <- mod$f0@\\
\mbox{}\verb@  spt <- round(mod$spt * N)@\\
\mbox{}\verb@  llik <- c(rowSums(fTiltMatrix * hp) %*% weights)@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtMacroRefIn\ \NWlink{nuweb3b}{3b}.

@@

\begin{flushleft} \small\label{scrap8}\raggedright\small
\NWtarget{nuweb6b}{} $\langle\,${\itshape Set starting values}\nobreak\ {\footnotesize {6b}}$\,\rangle\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@  N <- max(rowSums(Y))@\\
\mbox{}\verb@  nobs <- nrow(mm)@\\
\mbox{}\verb@  p <- ncol(mm)@\\
\mbox{}\verb@  @\\
\mbox{}\verb@  betas <- NULL@\\
\mbox{}\verb@  pooled <- CBData(data.frame(Trt = "All", NResp = Y[,1], ClusterSize = rowSums(Y), Freq=ceiling(weights)), @\\
\mbox{}\verb@                    trt="Trt", clustersize="ClusterSize", nresp="NResp")@\\
\mbox{}\verb@  est <- mc.est(pooled)@\\
\mbox{}\verb@  referencef0 <- est$Prob[est$ClusterSize == N]@\\
\mbox{}\verb@  # ensure all positive values@\\
\mbox{}\verb@  referencef0 <- (referencef0 + 1e-6)/(1+(N+1)*1e-6)@\\
\mbox{}\verb@  @\\
\mbox{}\verb@  fTiltMatrix <- matrix(rep(referencef0, times=nobs), byrow=TRUE, nrow=nobs, ncol=N+1)@\\
\mbox{}\verb@  spt <- 0:N@\\
\mbox{}\verb@@\\
\mbox{}\verb@  if (is.null(mu0))@\\
\mbox{}\verb@    mu0 <- weighted.mean(Y[,1]/rowSums(Y), weights)@\\
\mbox{}\verb@  @\\
\mbox{}\verb@  # hypergeometric terms for log-likelihood calculation@\\
\mbox{}\verb@  hp <- sapply(0:N, function(t)dhyper(x=Y[,1], m=rowSums(Y), n=N-rowSums(Y), k=t))   @\\
\mbox{}\verb@  @\\
\mbox{}\verb@  # initial log-likelihood@\\
\mbox{}\verb@  llik <- rowSums(fTiltMatrix * hp) %*% weights@\\
\mbox{}\verb@@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtMacroRefIn\ \NWlink{nuweb3b}{3b}.
